# Cluster analysis

Use divisive hierarchical clustering (DIANA).
See [Divisive Hierarchical Clustering Essentials](http://www.sthda.com/english/articles/28-hierarchical-clustering-essentials/94-divisive-hierarchical-clustering-essentials/).

```{r}
library(cluster)
library(ggdendro)
library(factoextra)
library(dendextend)
library(NbClust)
citation("factoextra")
pal <- lacroix_palette("PommeBaya", type = "discrete")
```

```{r}
metric <- "manhattan"
plotClusters <- function (df, title = "Cluster plot") {
  require(magrittr)
  require(ggplot2)
  require(factoextra)
  df %>% 
  fviz_cluster(data = mat, 
               palette = sprintf("%s7F", pal[1:k]), 
               geom = c("points"),
               labelsize = 8, 
               ellipse.type = "convex",
               star.plot = TRUE) + 
  ggtitle(title) +
  theme(panel.grid.minor = element_blank())
}
hcDIANA <- function(k) {
  hcut <-
    dist %>% 
    hcut(k = k, hc_func = "diana", hc_metric = metric)
  tab <- hcut$cluster %>% table()
  dend <- 
    hcut %>% 
    fviz_dend(k = k, 
              type = "circular",
              rect = TRUE, 
              rect_fill = TRUE,
              cex = 0.25, 
              k_colors = pal[1:k], 
              main = sprintf("DIANA\nk = %d", k))
  clust <- hcut %>% plotClusters(sprintf("DIANA\nDivisive coefficient = %.03f\nk = %d", diana$dc, k))
  silh <- 
    hcut %>% 
    fviz_silhouette(palette = sprintf("%s3F", pal[1:k]), ggtheme = theme_minimal()) + 
    theme(panel.grid.major.x = element_blank()) +
    ggtitle(sprintf("DIANA\nk = %d\nAverage silhouette width = %.03f\nCluster average silhouette widths = %s",
                    k,
                    hcut[["silinfo"]][["avg.width"]],
                    hcut[["silinfo"]][["clus.avg.widths"]] %>% round(3) %>% paste(collapse = ", ")))
  list(hcut = hcut,
       table = tab,
       dendrogram = dend,
       clusters2d = clust,
       silhouette = silh)
}
```

Use the **`r metric`** metric.

```{r}
# load("data/processed/dataframes.RData")
dfNonMissing <- 
  df %>% 
  na.omit()
dfXMetrics <- 
  dfNonMissing %>% 
  select(matches("ECBI|MAPS|SEPTI")) %>% 
  droplevels()
dfXParentChild <- 
  dfNonMissing %>% 
  select(matches("parent|child|birth", ignore.case = TRUE)) %>% 
  droplevels()
dfXDemog <- 
  dfNonMissing %>% 
  select(matches("lang|zip|community|distance|income|internet")) %>% 
  select(-matches("lang")) %>% 
  droplevels()
dfYPCB <- 
  dfNonMissing %>% 
  select(starts_with("PCB"))
```


## Cluster on PCB metrics

```{r clusterYPCB}
mat <- model.matrix(~ - 1 + ., data = dfYPCB) %>% scale()
dist <- mat %>% dist(method = metric)
dim(mat)
dim(dist)
colnames(mat)
hopkins <- get_clust_tendency(mat, n = nrow(mat) - 1, graph = FALSE)
diana <- dist %>% diana(metric = metric)
k <- 2
dianaYPCB <- hcDIANA(k)
dianaYPCB[["table"]]
dianaYPCB[["dendrogram"]]
dianaYPCB[["clusters2d"]]
dianaYPCB[["silhouette"]]
```

* Hopkins statistic is `r sprintf("%.03f", hopkins$hopkins_stat)`
* Analysis identified $k = `r k`$ clusters
* Divisive coefficient is `r sprintf("%.03f", diana$dc)`
* Average silhouette width is `r sprintf("%.03f", dianaXMetrics[["hcut"]][["silinfo"]][["avg.width"]])`


## Cluster on ECBI, MAPS, SEPTI metrics

```{r clusterXMetrics}
mat <- model.matrix(~ - 1 + ., data = dfXMetrics) %>% scale()
dist <- mat %>% dist(method = metric)
dim(mat)
dim(dist)
colnames(mat)
hopkins <- get_clust_tendency(mat, n = nrow(mat) - 1, graph = FALSE)
diana <- dist %>% diana(metric = metric)
k <- 3
dianaXMetrics <- hcDIANA(k)
dianaXMetrics[["table"]]
dianaXMetrics[["dendrogram"]]
dianaXMetrics[["clusters2d"]]
dianaXMetrics[["silhouette"]]
```

* Hopkins statistic is `r sprintf("%.03f", hopkins$hopkins_stat)`
* Analysis identified $k = `r k`$ clusters
* Divisive coefficient is `r sprintf("%.03f", diana$dc)`
* Average silhouette width is `r sprintf("%.03f", dianaXMetrics[["hcut"]][["silinfo"]][["avg.width"]])`


## Cluster on parent/child factors

```{r clusterXParentChild}
mat <- model.matrix(~ - 1 + ., data = dfXParentChild) %>% scale()
dist <- mat %>% dist(method = metric)
dim(mat)
dim(dist)
colnames(mat)
hopkins <- get_clust_tendency(mat, n = nrow(mat) - 1, graph = FALSE)
diana <- dist %>% diana(metric = metric)
k <- 4
dianaXParentChild <- hcDIANA(k)
dianaXParentChild[["table"]]
dianaXParentChild[["dendrogram"]]
dianaXParentChild[["clusters2d"]]
dianaXParentChild[["silhouette"]]
```

* Hopkins statistic is `r sprintf("%.03f", hopkins$hopkins_stat)`
* Analysis identified $k = `r k`$ clusters
* Divisive coefficient is `r sprintf("%.03f", diana$dc)`
* Average silhouette width is `r sprintf("%.03f", dianaXMetrics[["hcut"]][["silinfo"]][["avg.width"]])`


```{r}
bind_cols(dfXParentChild, 
          data.frame(cluster = dianaXParentChild[["hcut"]][["cluster"]] %>% factor())) %>%
  group_by(cluster, totalChildren, birthOrder) %>%
  summarize(n = n()) %>% 
  kable()
```


## Cluster on demographic factors

```{r clusterXDemog}
mat <- model.matrix(~ - 1 + ., data = dfXDemog) %>% scale()
dist <- mat %>% dist(method = metric)
dim(mat)
dim(dist)
colnames(mat)
hopkins <- get_clust_tendency(mat, n = nrow(mat) - 1, graph = FALSE)
diana <- dist %>% diana(metric = metric)
k <- 4
dianaXDemog <- hcDIANA(k)
dianaXDemog[["table"]]
dianaXDemog[["dendrogram"]]
dianaXDemog[["clusters2d"]]
dianaXDemog[["silhouette"]]
```

* Hopkins statistic is `r sprintf("%.03f", hopkins$hopkins_stat)`
* Analysis identified $k = `r k`$ clusters
* Divisive coefficient is `r sprintf("%.03f", diana$dc)`
* Average silhouette width is `r sprintf("%.03f", dianaXMetrics[["hcut"]][["silinfo"]][["avg.width"]])`

```{r}
bind_cols(dfXDemog, 
          data.frame(cluster = dianaXDemog[["hcut"]][["cluster"]] %>% factor())) %>%
  mutate(distanceLong = distance > 75) %>% 
  group_by(cluster, zipcodeClass, community, distanceLong) %>%
  summarize(n = n()) %>% 
  kable()
bind_cols(dfXDemog, 
          data.frame(cluster = dianaXDemog[["hcut"]][["cluster"]] %>% factor())) %>%
  group_by(cluster, zipcodeClass, income) %>%
  summarize(n = n()) %>% 
  kable()
bind_cols(dfXDemog, 
          data.frame(cluster = dianaXDemog[["hcut"]][["cluster"]] %>% factor())) %>%
  group_by(cluster, zipcodeClass, internet) %>%
  summarize(n = n()) %>% 
  kable()
```


## Save objects

```{r}
f <- "data/processed/clusterAnalysis.RData"
save(dianaYPCB, dianaXMetrics, dianaXParentChild, dianaXDemog, file = f)
file.info(f)
```
