# Preprocess data

Initial preprocesssing that needs to be done that is common to `PCB1_Total`, `PCB2_Tot`, and `PCB3_Total`.

```{r, echo = TRUE}
p <- 0.70
```

Split data set into `r sprintf("%.00f:%.00f", p * 100, (1 - p) * 100)` training:validation samples.

```{r, echo = TRUE}
inTrain <- createDataPartition(df$id, p = p)
dfTrain <- df[inTrain$Resample1, ]
dfValid <- df[-inTrain$Resample1, ]
```

Preprocess the training sample.

1. Exclude near-zero variance predictors
2. Impute missing values using k-nearest neighbor

```{r, echo = TRUE}
message(sprintf("Number of complete cases before imputation = %d",
                complete.cases(dfTrain) %>% sum()))
nzv <- 
  dfTrain %>% 
  select(-c(id, 
            PCB1_Total, PCB1_CondEmot, PCB1_DevHab, 
            PCB2_Tot, 
            PCB3_Total, PCB3_PCPonly, PCB3_Person, PCB3_Resource)) %>% 
  nearZeroVar(names = TRUE, saveMetric = TRUE) %>%
  mutate(varname = row.names(.)) %>% 
  filter(nzv == TRUE) %>% 
  select(varname, freqRatio, percentUnique, zeroVar, nzv) 
nzv %>% kable()
dfTrainPreProc1 <-
  dfTrain %>% 
  select(-one_of(nzv$varname))
dfOutcomes <- 
  dfTrainPreProc1 %>% 
  select(c(id,
           PCB1_Total, PCB1_CondEmot, PCB1_DevHab, 
           PCB2_Tot, 
           PCB3_Total, PCB3_PCPonly, PCB3_Person, PCB3_Resource))
dfTrainPreProc2 <- 
  dfTrainPreProc1 %>% 
  select(-c(id,
            PCB1_Total, PCB1_CondEmot, PCB1_DevHab, 
            PCB2_Tot, 
            PCB3_Total, PCB3_PCPonly, PCB3_Person, PCB3_Resource))
preProc <-
  dfTrainPreProc2 %>% 
  preProcess(method = c("nzv", "corr", "knnImpute"), verbose = TRUE)
preProc
dfTrainPreProc3 <- predict(preProc, dfTrainPreProc2)
dfTrainPreProc <- bind_cols(dfOutcomes, dfTrainPreProc3)
message(sprintf("Number of complete cases after imputation = %d",
                complete.cases(dfTrainPreProc) %>% sum()))
save(dfTrainPreProc, dfValid, dfTrain, npz, preProc, df, file = "data/processed/dataframes.RData")
rm(dfTrainPreProc1, dfTrainPreProc2, dfTrainPreProc3)
```

Set the control parameters.

```{r, echo = TRUE}
ctrl <- trainControl(method = "oob",
                     savePredictions = TRUE,
                     allowParallel = TRUE,
                     search = "random")
cores <- 24
```

Set the model and tuning parameter grid.

```{r, echo = TRUE}
library(earth)
citation("earth")
method <- "bagEarth"
modelLookup(method) %>% kable()
grid <- expand.grid(nprune = seq(10, 100, 10),
                    degree = seq(3))
grid %>% kable()
```
